---
title: "Locating tree species at SERC"
author: "Kelvin Acebron"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Uploading data
```{r}
ForestGEO_SERC <- read.csv("/Users/kelvinacebron/Documents/SERC NASA ROSES grant /03 Data/FORESTGEO SERC/SERC_censuses_with_UTMs_5_apr_2022.csv")
print(str(ForestGEO_SERC))
```
#### List all Families
```{r}
levels(factor(ForestGEO_SERC$Family))
```
There are 37 different treee families

#### List all Genus
```{r}
levels(factor(ForestGEO_SERC$Genus))
```
Theres a total of 53 different Genus

#### List all Species present in the data
```{r}
levels(factor(interaction(ForestGEO_SERC$Genus,ForestGEO_SERC$SpeciesName)))
```
Theres a total of 80 different tree species, with two unidentified Genus and unknown species

#### Now lets analyse the data based on available R package from ForestGEO website

Installation
Make sure your R environment is as follows:

R version is recent
All packages are updated (run update.packages(); maybe use ask = FALSE)
No other R session is running
Current R session is clean (click Session > Restart R)
Install the latest stable version of fgeo from CRAN with:

install.packages("fgeo")
Or install the development version of fgeo from GitHub with:

# install.packages("devtools")
devtools::install_github("forestgeo/fgeo.x")

```{r}
library(fgeo)
dff <- ForestGEO_SERC
sp.n <- length(unique(dff$Mnemonic))
dbh <- as.numeric(dff$DBH_1)
plot(dff$PX, dff$PY, col = heat.colors(sp.n)[factor(dff$Mnemonic)], cex = log(dbh/ 20), pch = 19)

sp.table <- sort(table(dff$Mnemonic), decreasing = TRUE)
barplot((sp.table))

dff.large <- subset(dff, DBH_1 >= 25)
sp.table <- sort(table(dff.large$Mnemonic), decreasing = TRUE)
barplot((sp.table))

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
